---
title: "Trends in provision of elective care in the independent sector"
author: "Alexander Lawless"
output: 
  html_document:
    theme: cerulean
   #toc: TRUE
   #toc_float: TRUE
   #toc_collapsed: TRUE
   #toc_depth: 4
   #number_sections: TRUE
    code_folding: hide
---

```{r setup, include=FALSE, warnings=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings=FALSE, message=FALSE, fig.width =12, fig.height = 9)

library(tidyverse)
library(janitor)
library(scales)
library(DT)
library(patchwork)

create_dt <- function(x) {
    
    DT::datatable(
        x
        , extensions = "Buttons"
        , options = list(
            dom = "Blfrtip"
            , buttons = c("copy", "csv")
            , lengthMenu = list(
                c(10, 25, 50, -1)
                , c(10, 25, 50, "All"))))
  }

# Set SU theme ####
SU_colours <- c (
  `orange`                     = grDevices::rgb(248,191,7, maxColorValue = 255),# "#f9bf07",
  `charcoal`                   = grDevices::rgb(44,40,37, maxColorValue = 255),# "#2c2825",
  `slate`                      = grDevices::rgb(104,111,115, maxColorValue = 255), # "#686f73",
  `blue`                       = grDevices::rgb(88,29,193, maxColorValue = 255), # "#5881c1",
  `red`                        = grDevices::rgb(236,101,85, maxColorValue = 255), # "#ec6555",
  #additional accent colours from word doc template
  `yellow`                     = grDevices::rgb(252,229,155, maxColorValue = 255),
  `grey`                       = grDevices::rgb(163,168,172, maxColorValue = 255),
  `white`                      = grDevices::rgb(255,255,255, maxColorValue = 255),
  #light and dark ends from colour theme in word doc
  `light orange`               = grDevices::rgb(253,242,205, maxColorValue = 255),
  `dark orange`                = grDevices::rgb(124,95,3, maxColorValue = 255),
  `light charcoal`             = grDevices::rgb(235,233,231, maxColorValue = 255),
  `dark charcoal`              = 	"#000000",#black
  `light slate`                = grDevices::rgb(224,226,227, maxColorValue = 255),
  `dark slate`                 = grDevices::rgb(51,55,57, maxColorValue = 255),
  `light blue`                 = grDevices::rgb(221,229,242, maxColorValue = 255),
  `dark blue`                  = grDevices::rgb(38,61,102, maxColorValue = 255),
  `light red`                  = grDevices::rgb(251,224,220, maxColorValue = 255),
  `dark red`                   = grDevices::rgb(144,29,16, maxColorValue = 255),
  `light yellow`               = grDevices::rgb(254,249,235, maxColorValue = 255),
  `dark yellow`                = grDevices::rgb(197,152,5, maxColorValue = 255),
  `light grey`                 = grDevices::rgb(236,237,238, maxColorValue = 255),
  `dark grey`                  = grDevices::rgb(79,84,88, maxColorValue = 255),
  `light white`                = grDevices::rgb(242,242,242, maxColorValue = 255),
  `dark white`                 = grDevices::rgb(127,127,127, maxColorValue = 255),
  `red2`                       = grDevices::rgb(215,25,28, maxColorValue = 255),
  `orange2`                    = grDevices::rgb(253,174,97, maxColorValue = 255),
  `yellow2`                    = grDevices::rgb(255,255,191, maxColorValue = 255),
  `green2`                     = grDevices::rgb(171,221,164, maxColorValue = 255),
  `blue2`                      = grDevices::rgb(43,131,186, maxColorValue = 255) #"#2b83ba"
)

SU_cols <- function(...) {
  cols <- c(...)
  
  if (is.null(cols))
    return (SU_colours)
  
  SU_colours[cols]
}

SU_palettes <- list(
  `main` = SU_cols("orange","charcoal","slate","blue","red"),
  `oranges` = SU_cols("light orange","orange","dark orange"),
  `slates` = SU_cols("light slate","slate","dark slate"),
  `mixed` = SU_cols("dark red","orange","yellow","light blue","slate"),
  `oj_coal` = SU_cols("yellow","orange","red","dark red","dark charcoal"),
  `oj_red` = SU_cols("yellow","orange","red","dark red"),
  `white_oj_coal` = SU_cols("white","yellow","orange","red","dark red","dark charcoal"),#added since shared
  `lyellow_oj_coal` = SU_cols("light yellow","orange","red","dark red","dark charcoal"),#added since shared
  `wy_oj_coal` = SU_cols("white","light yellow","yellow","orange","red","dark red","charcoal","dark charcoal"),
  `red_coal` = SU_cols("red","dark red","charcoal","dark charcoal"),
  `blue_yellow_red` = SU_cols("red2","orange2","yellow2","green2","blue2"),
  `red_yellow_blue` = SU_cols("blue2","green2","yellow2","orange2","red2")
)


SU_pal <- function(palette = "main", reverse = FALSE, ...) {
  pal <- SU_palettes[[palette]]
  
  if (reverse) pal <- rev(pal)
  
  colorRampPalette(pal, ...)
}


scale_color_SU <- function(palette = "main", discrete = TRUE, reverse = FALSE, ...) {
  pal <- SU_pal(palette = palette, reverse = reverse)
  
  if (discrete) {
    discrete_scale("colour", paste0("SU_", palette), palette = pal, ...)
  } else {
    scale_color_gradientn(colours = pal(256), ...)
  }
}

scale_fill_SU <- function(palette = "main", discrete = TRUE, reverse = FALSE, ...) {
  pal <- SU_pal(palette = palette, reverse = reverse)
  
  if (discrete) {
    discrete_scale("fill", paste0("SU_", palette), palette = pal, ...)
  } else {
    scale_fill_gradientn(colours = pal(256), ...)
  }
}  

theme_SU <-   function (base_size){
  theme_minimal(
    #base_family = "Segoe UI", 
    base_size=12
  ) %+replace% 
    theme(axis.title = element_text(size=11, face="bold",colour=SU_cols("charcoal")),
          plot.title = element_text(hjust=0,face="bold",size=12,colour=SU_cols("charcoal"),margin=margin(b=4,unit="pt")),
          plot.subtitle = element_text(hjust=0,face="italic",size=10,colour=SU_cols("charcoal"),margin=margin(b=4,unit="pt")),
          plot.caption = element_text(hjust = 0,face="italic",size=9,colour=SU_cols("slate"),margin=margin(b=4,unit="pt")),
          legend.text = element_text(size=10,colour=SU_cols("charcoal")),
          legend.title = element_text(face="bold",size=11,colour=SU_cols("charcoal"),margin=margin(b=4,unit="pt")))
}

theme_set(theme_SU())

# Import data/ functions ----

hrg_lookup <-
  tribble(
  ~hrg_3, ~procedure_desc, ~procedure_desc_short, ~procedure_group,
  'BZ3', "Cataract and Lens Procedures",       "Cataract",            "Subchapter BZ – Eyes and Periorbita Procedures and Disorders",
  'BZ4', "Oculoplastics Procedures",           "Oculoplastics",       "Subchapter BZ – Eyes and Periorbita Procedures and Disorders",
  'BZ5', "Orbit and Lacrimal Procedures",      "Orbit and Lacrimal",  "Subchapter BZ – Eyes and Periorbita Procedures and Disorders",
  'BZ6', "Cornea and Sclera Procedures",       "Cornea and Sclera",   "Subchapter BZ – Eyes and Periorbita Procedures and Disorders",
  'BZ7', "Ocular Motility Procedures",         "Ocular Motility",     "Subchapter BZ – Eyes and Periorbita Procedures and Disorders",
  'BZ8', "Vitreous Retinal Procedures",        "Vitreous Retinal",    "Subchapter BZ – Eyes and Periorbita Procedures and Disorders",
  'BZ9', "Glaucoma Procedures",                "Glaucoma",            "Subchapter BZ – Eyes and Periorbita Procedures and Disorders",
  
  'HN1', "Hip Procedures for Non-Trauma",      "Hip",                 "Subchapter HN – Orthopaedic Non-Trauma Procedures",
  'HN2', "Knee Procedures for Non-Trauma",     "Knee",                "Subchapter HN – Orthopaedic Non-Trauma Procedures",
  'HN3', "Foot Procedures for Non-Trauma",     "Foot",                "Subchapter HN – Orthopaedic Non-Trauma Procedures",
  'HN4', "Hand Procedures for Non-Trauma",     "Hand",                "Subchapter HN – Orthopaedic Non-Trauma Procedures",
  'HN5', "Shoulder Procedures for Non-Trauma", "Shoulder",            "Subchapter HN – Orthopaedic Non-Trauma Procedures", 
  'HN6', "Elbow Procedures for Non-Trauma",    "Elbow",               "Subchapter HN – Orthopaedic Non-Trauma Procedures"
  )

national_data <- 
  read_csv("national_data.csv") %>% 
  mutate(imd_quintile = 
           case_when(imd_decile %in% c(1,2) ~ 1,
                     imd_decile %in% c(3,4) ~ 2,
                     imd_decile %in% c(5,6) ~ 3,
                     imd_decile %in% c(7,8) ~ 4,
                     imd_decile %in% c(9,10) ~ 5)) %>% 
  mutate(duration_elective_wait_range = 
           factor(duration_elective_wait_range, 
                  levels = 
                    c("0-50","50-100","100-150","150-200","200-250","250-300", 
                    "300-350","350-400","400-450","450-500","500+")))

national_data_ortho <- 
  national_data %>% 
  filter(speciality == "Orthopaedic")

national_data_ophthal <- 
  national_data %>% 
  filter(speciality == "Ophthalmology")

sum_spells_function <- function(data) {
  
  data %>%  
    summarise(n_spells_IP = sum(n_spells_IP, na.rm = TRUE),
              n_spells_OP = sum(n_spells_OP, na.rm = TRUE),
              cost = sum(cost_IP, cost_OP, na.rm = TRUE)
              )
  
}


# National trends by procedure
national_data_procedure <-
  national_data %>% 
  group_by(der_activity_month, type, speciality, procedure_desc_short) %>% 
  summarise(n_spells_IP = sum(n_spells_IP, na.rm = TRUE),
            n_spells_OP = sum(n_spells_OP, na.rm = TRUE)) %>% 
  drop_na(procedure_desc_short)


# National trends by demographic
age_range_order <-
  tribble(
    ~age_order, ~age_range,   
    1,  "0-9",    
    2,  "10-19",    
    3,  "20-29",    
    4,  "30-39",    
    5,  "40-49",    
    6,  "50-59",    
    7,  "60-69",    
    8,  "70-79",    
    9,  "80-89",   
    10, "90-99",    
    11, "100+"
    )

ethnicity_lookup <- read_csv("ethnicity_lookup.csv")

wrangle_function <- function(data, var_1) {
  
  data <- 
    data %>% 
    mutate(var_1 = {{var_1}})
  
  data %>%   
    group_by(der_activity_month, var_1, type) %>% 
    sum_spells_function(.) %>% 
    group_by(der_activity_month, type, var_1) %>% 
    mutate(total_activity = sum(n_spells_IP, n_spells_OP)) %>% 
    pivot_longer(cols = c(-der_activity_month, -type, -var_1)) %>% 
    pivot_wider(id_cols = c(der_activity_month, var_1, name), 
                names_from = type,
                values_from = value
                ) %>% 
    mutate(`Independent Sector` = case_when(is.na(`Independent Sector`) ~ 0, TRUE ~ `Independent Sector`)) %>% 
    group_by(der_activity_month, var_1, name) %>% 
    mutate(prop = `Independent Sector`/ sum(`Independent Sector`, `NHS`) * 100) %>% 
    mutate(name = case_when(str_detect(name, "IP") ~ "Inpatient admissions", 
                            str_detect(name, "OP") ~ "Outpatient appointments",
                            str_detect(name, "cost") ~ "Costs",
                            TRUE ~ "All activity")) %>% 
    filter(der_activity_month > "2018-01-01" &
             der_activity_month < "2022-11-01") %>%
    filter(name != "All activity") %>% 
    ungroup() %>% 
    filter(name != "Costs")
}

graph_function <- function(data, subtitle_speciality ) {
  
  data %>% 
    ggplot(aes(x = der_activity_month, y = prop, colour = name)) +
    geom_smooth(method = "loess", span = 0.3) +
    #facet_grid(imd_quintile~ethnicity_broad) +
    scale_color_SU() +
    scale_y_continuous(labels = comma, oob = squish) +
    theme(strip.background = element_rect(fill = NA, colour = "grey"),
          strip.text = element_text(face = "bold"), 
          legend.position = "bottom",
          axis.text.x = element_text(angle = 90),
          axis.title.x = element_blank()
          ) +
    labs(y = "Independent sector proportion (%)",
         colour = "",
         title = "Independent sector proportion of activity",
         subtitle = paste0(subtitle_speciality, " elective procedures | National | 2018-22")
    )
}

# Local/regional data 
# Look up(s)
stp_region_lookup <- 
  read_csv("STP to NHS England Region (April 2020).csv") %>% 
  clean_names()

# Data
stp_data <- 
  read_csv("stp_data.csv") %>% 
  left_join(stp_region_lookup, by = c("stp21nm" = "stp20nm")) %>% 
  mutate(nhser20nm = case_when(str_detect(stp21nm, "Sussex")~ "South East", TRUE ~ nhser20nm))

stp_data_ortho <-
  stp_data %>% 
  filter(speciality == "Orthopaedic")

stp_data_ophthal <-
  stp_data %>% 
  filter(speciality == "Ophthalmology")

stp_name_short_lookup <-
  tribble(
  ~stp_name, ~stp_name_short,
  
  "Bath and North East Somerset, Swindon and Wiltshire",  "Bath and North East Somerset, Swindon and Wiltshire",
  "Bedfordshire, Luton and Milton Keynes", "Bedfordshire, Luton and Milton Keynes",
  "Birmingham and Solihull", "Birmingham and Solihull",
  "Bristol, North Somerset and South Gloucestershire", "Bristol, North Somerset and South Gloucestershire",
  "Buckinghamshire, Oxfordshire and Berkshire West", "Buckinghamshire, Oxfordshire and Berkshire West",
  "Cambridgeshire and Peterborough", "Cambridgeshire and Peterborough",
  "Cheshire and Merseyside", "Cheshire and Merseyside",
  "Cornwall and the Isles of Scilly Health and Social Care Partnership", "Cornwall and the Isles of Scilly",
  "Coventry and Warwickshire", "Coventry and Warwickshire",
  "Cumbria and North East", "Cumbria and North East",
  "Devon", "Devon",
  "Dorset", "Dorset",
  "East London Health and Care Partnership", "East London",
  "Frimley Health and Care ICS", "Frimley ",
  "Gloucestershire", "Gloucestershire",
  "Greater Manchester Health and Social Care Partnership", "Greater Manchester",
  "Hampshire and the Isle of Wight", "Hampshire and the Isle of Wight",
  "Healthier Lancashire and South Cumbria", "Healthier Lancashire and South Cumbria",
  "Herefordshire and Worcestershire", "Herefordshire and Worcestershire",
  "Hertfordshire and West Essex", "Hertfordshire and West Essex",
  "Humber, Coast and Vale", "Humber, Coast and Vale",
  "Joined Up Care Derbyshire", "Joined Up Care Derbyshire",
  "Kent and Medway", "Kent and Medway",
  "Leicester, Leicestershire and Rutland", "Leicester, Leicestershire and Rutland",
  "Lincolnshire", "Lincolnshire",
  "Mid and South Essex", "Mid and South Essex",
  "Norfolk and Waveney Health and Care Partnership", "Norfolk and Waveney",
  "North London Partners in Health and Care", "North London",
  "North West London Health and Care Partnership", "North West London",
  "Northamptonshire", "Northamptonshire",
  "Nottingham and Nottinghamshire Health and Care", "Nottingham and Nottinghamshire",
  "Our Healthier South East London", "South East London",
  "Shropshire and Telford and Wrekin", "Shropshire and Telford and Wrekin",
  "Somerset", "Somerset",
  "South West London Health and Care Partnership", "South West London",
  "South Yorkshire and Bassetlaw", "South Yorkshire and Bassetlaw",
  "Staffordshire and Stoke on Trent", "Staffordshire and Stoke on Trent",
  "Suffolk and North East Essex", "Suffolk and North East Essex",
  "Surrey Heartlands Health and Care Partnership", "Surrey Heartlands",
  "Sussex Health and Care Partnership", "Sussex",
  "The Black Country and West Birmingham", "The Black Country and West Birmingham",
  "West Yorkshire and Harrogate (Health and Care Partnership)", "West Yorkshire and Harrogate"
  )

ICB_graph_function <- function(data_input, speciality) {
  
  data <- data_input
  
  function_1 <- function(region){
    data %>% 
      wrangle_function(., stp21nm) %>% 
      left_join(stp_region_lookup, by = c("var_1" = "stp20nm")) %>% 
      mutate(nhser20nm = case_when(str_detect(var_1, "Sussex")~ "South East", TRUE ~ nhser20nm)) %>% 
      filter(nhser20nm == region) %>% 
      filter(!is.nan(prop)) %>% 
      left_join(stp_name_short_lookup, by = c("var_1" = "stp_name")) %>% 
      mutate(stp_name_short = case_when(nchar(stp_name_short) <= 20 ~ stp_name_short, 
                                     TRUE ~ paste0(stringr::str_sub(stp_name_short, 0, 20), "..."))
             ) %>%
      
      ggplot(aes(x = der_activity_month, y = prop, colour = name)) +
      geom_smooth(method = "loess", span = 0.3, se = FALSE) +
      facet_grid(str_wrap(nhser20nm, 12) ~ str_wrap(stp_name_short, 15)) +
      scale_color_SU() +
      scale_y_continuous(limits = c(0,100), oob = squish) +
      scale_x_date(date_breaks = "2 years", labels = label_date_short()) +
      theme(strip.background = element_rect(fill = NA, colour = "grey"),
            strip.text.y = element_text(face = "bold", angle = 0, size = 8), 
            strip.text.x = element_text(size = 7), 
            axis.text.x = element_text(angle = 90),
            legend.position = "none",
            plot.title = element_blank(),
            plot.subtitle = element_blank(),
            axis.title = element_blank())
  }  
  
  plot_1 <- function_1("East of England")
  plot_2 <- function_1("London")
  plot_3 <- function_1("Midlands")
  plot_4 <- function_1("North East and Yorkshire")
  plot_5 <- function_1("North West")
  plot_6 <- function_1("South East")
  plot_7 <- function_1("South West") +
    theme(legend.position = "bottom") + 
    labs(colour = "")
  
  patch_1 <-
    plot_3 +
    (plot_1 + 
       plot_2 +
       plot_4 +
       plot_5 +
       plot_6 +
       plot_7 +
       plot_layout(ncol = 2)) +
    plot_layout(nrow = 2, 
                heights = c(0.2, 0.8)) +    
    plot_annotation(title = "Independent sector provision share by ICB and region", 
                    subtitle = paste0(speciality, " elective care | 2018-22")
                    )
  
  patch_1
  
  }

```
<hr style="border-top: 2px solid #f9bf07">
# Introduction

The role of the independent sector in the provision of elective care has been brought into the spotlight by rising pressures of the NHS and growing waiting lists. Our analysis aims to contribute to existing work undertaken by the Health Foundation investigating the landscape of [independent sector healthcare services](https://www.health.org.uk/publications/long-reads/waiting-for-nhs-hospital-care-the-role-of-the-independent-sector). 

The Health Foundation found an increasing proportion of care is delivered by the independent sector; this varies by specialitiy and is highest in T&O, GI and ophthalmology. Treatment rates in all specialities are not achieving the 30%-increase target set to address the backlog of care. The Health Foundation suggests uptake in the independent sector is matching the reductions in NHS provided care. It was also found that the independent sector share of provision varies regionally and by deprivation measures. 

We aim to further the above analysis by utilising patient-level inpatient and outpatient data to more closely examine trends in independent sector care. Given the above findings, we will focus on ophthalmology and orthopaedic care, presenting our findings by demographics, geographically and by procedure. 

<hr style="border-top: 2px solid #f9bf07">

# Methods

In an exploratory fashion, we will present the changing market share of independent care in elective ophthalmic and orthopaedic services. We will be guided by our preliminary findings starting by assessing patterns in national data developing on to consider local findings.   

## Data collection 

<div class = "row">
<div class = "col-md-5">
<br>
We gathered acute inpatient and outpatient data via the Secondary Uses Services (SUS) dataset, accessed through the National Commissioning Data Repository (NCDR) platform. To capture the pre-pandemic context, we limit our data to the years 2017 through to 2022. We count completed inpatient admissions or outpatient attendances filtered to only include non-emergency contacts with selected Healthcare Resource Group (HRG) codes. 

The HRG code describes the content of an admission in terms of resource required and is linked to the of commissioning care. We use commissioning codes rather than procedure codes because data completeness is preferable and they offer the suitable level of granularity for data processing and presentation of findings. All HRG codes were updated to their corresponding code in the most recent HRG version. The activity we captured was linked to the following HRG codes (right).

Data was accessed at a national level with local feilds included to indicate the patient's ICS and local authority of residence. Similarly, the low level geography (LSOA) of the patient was included to link to the Index of Multiple Deprivation (IMD) to assign each admission to a deprivation decile (1 = most deprived 10% of the population, 10 = the most affluent 10% of the population).

The SUS data contains only NHS-funded care. We identified the sector in which the care was delivered by assessing the ODS provider type, linked to site (hospital/care centre) in which the care occurred. We compare the rate of NHS-funded care delivered in NHS sites to NHS-funded care delivered in independent sector sites. Limitations of our data dictate that we do not count any privately funded care, regardless of site of delivered.

</div>
  
<div class = "col-md-7">
<br>
***Subchapter BZ – Eyes and Periorbita Procedures and Disorders:***
```{r hrg_codes_eyes_2}
create_dt(
  hrg_lookup %>% 
  filter(str_detect(procedure_group, "Eyes")) %>% 
  select(1,2) %>% 
  rename(HRG = hrg_3, 
         Procedure = procedure_desc)
  )
```

***Subchapter HN – Orthopaedic Non-Trauma Procedures:***
```{r hrg_codes_ortho_2}
create_dt(
  hrg_lookup %>% 
  filter(str_detect(procedure_group, "Orthopaedic")) %>% 
  select(1,2) %>% 
  rename(HRG = hrg_3, 
         Procedure = procedure_desc)
  )
```

</div>
</div>

## Data processing 

Activity data was grouped by demographic, geographic and activity-type variables and counts of episodes, spells (admissions or attendances) and individuals were calculated. Aggregated data was exported from the NCDR SQL server and processed locally in RStudio/RMarkdown. 

For each group of interest we had activity counts for care delivered in both the independent sector (IS) and in NHS clinical settings. We calculated the independent sector market share (%) as follows:

$$
IndependentSectorShare  = ISActivity / (ISActivity + NHSActivity) * 100
$$

Incomplete admissions/attendances, i.e. those without a discharge date, were excluded. Similarly, those without valid identification numbers were excluded.

Upon publication of our findings, the underlying code will be available on [Github](https://github.com/alexanderlawless).

<hr style="border-top: 2px solid #f9bf07">

# Results

National results link: 

Orthopaedic Sub-national results: 

Ophthalmic Sub-national results: 


## Local trends

### Region

#### Volume {.tabset}

##### Orthopaedic
```{r}
stp_data_ortho %>% 
  wrangle_function(., nhser20nm) %>% 
  filter(!is.nan(prop)) %>% 
  pivot_longer(cols = c(`Independent Sector`, NHS),
               names_to = "sector") %>% 
  ggplot(aes(x = der_activity_month, y = value, colour = var_1)) +
  geom_smooth(method = 'loess', span = 0.4, se = FALSE, size = 1.25) +
  facet_grid(name~sector) +
  scale_color_SU() +
  scale_y_continuous(labels = comma) +
  theme(strip.text = element_text(face = "bold"),
        strip.background = element_rect(fill = NA, colour = "grey"),
        axis.title.x = element_blank()
        ) +
  labs(y = "Activity", 
       colour = "NHSE region",
       title = "Regional activity volumes by care type and sector: Orthopaedic",
       subtitle = "Monthly elective activity counts | NHSE region | 2018-22")
```

<div class = "row">
<div class = "col-md-1">
</div>
  
<div class = "col-md-10">

Graph desc... 

</div>

<div class = "col-md-1">
</div>

</div>

##### Ophthalmology
```{r}
stp_data_ophthal %>% 
  wrangle_function(., nhser20nm) %>% 
  filter(!is.nan(prop)) %>% 
  pivot_longer(cols = c(`Independent Sector`, NHS),
               names_to = "sector") %>% 
  ggplot(aes(x = der_activity_month, y = value, colour = var_1)) +
  geom_smooth(method = 'loess', span = 0.4, se = FALSE, size = 1.25) +
  facet_grid(name~sector, scales = "free") +
  scale_color_SU() +
  scale_y_continuous(labels = comma) +
  theme(strip.text = element_text(face = "bold"),
        strip.background = element_rect(fill = NA, colour = "grey"),
        axis.title.x = element_blank()
        ) +
  labs(y = "Activity", 
       colour = "NHSE region",
       title = "Regional activity volumes by care type and sector: Ophthalmology",
       subtitle = "Monthly elective activity counts | NHSE region | 2018-22")
```

<div class = "row">
<div class = "col-md-1">
</div>
  
<div class = "col-md-10">

Graph desc... 

</div>

<div class = "col-md-1">
</div>

</div>

## {.unlisted .unnumbered}

Volume section desc

#### Independent sector share {.tabset}

##### Orthopaedic
```{r}
stp_data_ortho %>% 
  wrangle_function(., nhser20nm) %>% 
  filter(!is.nan(prop)) %>% 
  graph_function(., "Orthopaedic") +
  facet_grid(~str_wrap(var_1,15))
```

<div class = "row">
<div class = "col-md-1">
</div>
  
<div class = "col-md-10">

Graph desc... 

</div>

<div class = "col-md-1">
</div>

</div>

##### Ophthalmology
```{r}
stp_data_ophthal %>% 
  wrangle_function(., nhser20nm) %>% 
  filter(!is.nan(prop)) %>% 
  graph_function(., "Ophthalmology") +
  facet_grid(~str_wrap(var_1,15))
```

<div class = "row">
<div class = "col-md-1">
</div>
  
<div class = "col-md-10">

Graph desc... 

</div>

<div class = "col-md-1">
</div>

</div>

## {.unlisted .unnumbered}

Independent sector share section desc ...

### STP/ICB

#### Speciality {.tabset}

##### Orthopaedic 
```{r}
ICB_graph_function(stp_data_ortho, "Orthopaedic")
```


<div class = "row">
<div class = "col-md-1">
</div>
  
<div class = "col-md-10">

Graph desc... 

</div>

<div class = "col-md-1">
</div>

</div>

##### Ophthalmology 
```{r}
ICB_graph_function(stp_data_ophthal, "Ophthalmology")
```



<div class = "row">
<div class = "col-md-1">
</div>
  
<div class = "col-md-10">

Graph desc... 

</div>

<div class = "col-md-1">
</div>

</div>

## {.unlisted .unnumbered}

STP/ICB section desc ...

#### Procedure {.tabset}

##### Cataract 
```{r}
ICB_graph_function(stp_data_ophthal %>% 
                     filter(procedure_desc_short == "Cataract"), 
                   "Ophthalmology - Cataract")
```


<div class = "row">
<div class = "col-md-1">
</div>
  
<div class = "col-md-10">

Graph desc... 

</div>

<div class = "col-md-1">
</div>

</div>

##### Vitreous Retinal 
```{r}
ICB_graph_function(stp_data_ophthal %>% 
                     filter(procedure_desc_short == "Vitreous Retinal"), 
                   "Ophthalmology - Vitreous Retinal")
```



<div class = "row">
<div class = "col-md-1">
</div>
  
<div class = "col-md-10">

Graph desc... 

</div>

<div class = "col-md-1">
</div>

</div>

##### Glaucoma 
```{r}
ICB_graph_function(stp_data_ophthal %>% 
                     filter(procedure_desc_short == "Glaucoma"), 
                   "Ophthalmology - Glaucoma")
```



<div class = "row">
<div class = "col-md-1">
</div>
  
<div class = "col-md-10">

Graph desc... 

</div>

<div class = "col-md-1">
</div>

</div>

## {.unlisted .unnumbered}

Section desc 

#### Demographic - deprivation {.tabset}

##### Quintile 1 
```{r}
ICB_graph_function(stp_data_ophthal %>% 
                     filter(imd_quintile == 1), 
                   "Ophthalmology - IMD quintile 1")
```


<div class = "row">
<div class = "col-md-1">
</div>
  
<div class = "col-md-10">

Graph desc... 

</div>

<div class = "col-md-1">
</div>

</div>

##### 2 
```{r}
ICB_graph_function(stp_data_ophthal %>% 
                     filter(imd_quintile == 2), 
                   "Ophthalmology - IMD quintile 2")
```


<div class = "row">
<div class = "col-md-1">
</div>
  
<div class = "col-md-10">

Graph desc... 

</div>

<div class = "col-md-1">
</div>

</div>

##### 3 
```{r}
ICB_graph_function(stp_data_ophthal %>% 
                     filter(imd_quintile == 3), 
                   "Ophthalmology - IMD quintile 3")
```


<div class = "row">
<div class = "col-md-1">
</div>
  
<div class = "col-md-10">

Graph desc... 

</div>

<div class = "col-md-1">
</div>

</div>

##### 4 
```{r}
ICB_graph_function(stp_data_ophthal %>% 
                     filter(imd_quintile == 4), 
                   "Ophthalmology - IMD quintile 4")
```


<div class = "row">
<div class = "col-md-1">
</div>
  
<div class = "col-md-10">

Graph desc... 

</div>

<div class = "col-md-1">
</div>

</div>

##### Quintile 5 
```{r}
ICB_graph_function(stp_data_ophthal %>% 
                     filter(imd_quintile == 5), 
                   "Ophthalmology - IMD quintile 5")
```


<div class = "row">
<div class = "col-md-1">
</div>
  
<div class = "col-md-10">

Graph desc... 

</div>

<div class = "col-md-1">
</div>

</div>

## {.unlisted .unnumbered}

Section desc 


<hr style="border-top: 2px solid #f9bf07">




# Further discussion/Questions

* ...

