---
title: "Independent sector trends in provision of elective care"
author: "Alexander Lawless"
output: 
  html_document:
    theme: cerulean
    toc: TRUE
    toc_float: TRUE
    toc_collapsed: TRUE
    toc_depth: 4
    number_sections: TRUE
    code_folding: hide
---

```{r setup, include=FALSE, warnings=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings=FALSE, message=FALSE, fig.width =12, fig.height = 9)

library(tidyverse)
library(janitor)
library(scales)
library(DT)
library(patchwork)

create_dt <- function(x) {
    
    DT::datatable(
        x
        , extensions = "Buttons"
        , options = list(
            dom = "Blfrtip"
            , buttons = c("copy", "csv")
            , lengthMenu = list(
                c(10, 25, 50, -1)
                , c(10, 25, 50, "All"))))
  }

# Set SU theme ####
SU_colours <- c (
  `orange`                     = grDevices::rgb(248,191,7, maxColorValue = 255),# "#f9bf07",
  `charcoal`                   = grDevices::rgb(44,40,37, maxColorValue = 255),# "#2c2825",
  `slate`                      = grDevices::rgb(104,111,115, maxColorValue = 255), # "#686f73",
  `blue`                       = grDevices::rgb(88,29,193, maxColorValue = 255), # "#5881c1",
  `red`                        = grDevices::rgb(236,101,85, maxColorValue = 255), # "#ec6555",
  #additional accent colours from word doc template
  `yellow`                     = grDevices::rgb(252,229,155, maxColorValue = 255),
  `grey`                       = grDevices::rgb(163,168,172, maxColorValue = 255),
  `white`                      = grDevices::rgb(255,255,255, maxColorValue = 255),
  #light and dark ends from colour theme in word doc
  `light orange`               = grDevices::rgb(253,242,205, maxColorValue = 255),
  `dark orange`                = grDevices::rgb(124,95,3, maxColorValue = 255),
  `light charcoal`             = grDevices::rgb(235,233,231, maxColorValue = 255),
  `dark charcoal`              = 	"#000000",#black
  `light slate`                = grDevices::rgb(224,226,227, maxColorValue = 255),
  `dark slate`                 = grDevices::rgb(51,55,57, maxColorValue = 255),
  `light blue`                 = grDevices::rgb(221,229,242, maxColorValue = 255),
  `dark blue`                  = grDevices::rgb(38,61,102, maxColorValue = 255),
  `light red`                  = grDevices::rgb(251,224,220, maxColorValue = 255),
  `dark red`                   = grDevices::rgb(144,29,16, maxColorValue = 255),
  `light yellow`               = grDevices::rgb(254,249,235, maxColorValue = 255),
  `dark yellow`                = grDevices::rgb(197,152,5, maxColorValue = 255),
  `light grey`                 = grDevices::rgb(236,237,238, maxColorValue = 255),
  `dark grey`                  = grDevices::rgb(79,84,88, maxColorValue = 255),
  `light white`                = grDevices::rgb(242,242,242, maxColorValue = 255),
  `dark white`                 = grDevices::rgb(127,127,127, maxColorValue = 255),
  `red2`                       = grDevices::rgb(215,25,28, maxColorValue = 255),
  `orange2`                    = grDevices::rgb(253,174,97, maxColorValue = 255),
  `yellow2`                    = grDevices::rgb(255,255,191, maxColorValue = 255),
  `green2`                     = grDevices::rgb(171,221,164, maxColorValue = 255),
  `blue2`                      = grDevices::rgb(43,131,186, maxColorValue = 255) #"#2b83ba"
)

SU_cols <- function(...) {
  cols <- c(...)
  
  if (is.null(cols))
    return (SU_colours)
  
  SU_colours[cols]
}

SU_palettes <- list(
  `main` = SU_cols("orange","charcoal","slate","blue","red"),
  `oranges` = SU_cols("light orange","orange","dark orange"),
  `slates` = SU_cols("light slate","slate","dark slate"),
  `mixed` = SU_cols("dark red","orange","yellow","light blue","slate"),
  `oj_coal` = SU_cols("yellow","orange","red","dark red","dark charcoal"),
  `oj_red` = SU_cols("yellow","orange","red","dark red"),
  `white_oj_coal` = SU_cols("white","yellow","orange","red","dark red","dark charcoal"),#added since shared
  `lyellow_oj_coal` = SU_cols("light yellow","orange","red","dark red","dark charcoal"),#added since shared
  `wy_oj_coal` = SU_cols("white","light yellow","yellow","orange","red","dark red","charcoal","dark charcoal"),
  `red_coal` = SU_cols("red","dark red","charcoal","dark charcoal"),
  `blue_yellow_red` = SU_cols("red2","orange2","yellow2","green2","blue2"),
  `red_yellow_blue` = SU_cols("blue2","green2","yellow2","orange2","red2")
)


SU_pal <- function(palette = "main", reverse = FALSE, ...) {
  pal <- SU_palettes[[palette]]
  
  if (reverse) pal <- rev(pal)
  
  colorRampPalette(pal, ...)
}


scale_color_SU <- function(palette = "main", discrete = TRUE, reverse = FALSE, ...) {
  pal <- SU_pal(palette = palette, reverse = reverse)
  
  if (discrete) {
    discrete_scale("colour", paste0("SU_", palette), palette = pal, ...)
  } else {
    scale_color_gradientn(colours = pal(256), ...)
  }
}

scale_fill_SU <- function(palette = "main", discrete = TRUE, reverse = FALSE, ...) {
  pal <- SU_pal(palette = palette, reverse = reverse)
  
  if (discrete) {
    discrete_scale("fill", paste0("SU_", palette), palette = pal, ...)
  } else {
    scale_fill_gradientn(colours = pal(256), ...)
  }
}  

theme_SU <-   function (base_size){
  theme_minimal(
    #base_family = "Segoe UI", 
    base_size=12
  ) %+replace% 
    theme(axis.title = element_text(size=11, face="bold",colour=SU_cols("charcoal")),
          plot.title = element_text(hjust=0,face="bold",size=12,colour=SU_cols("charcoal"),margin=margin(b=4,unit="pt")),
          plot.subtitle = element_text(hjust=0,face="italic",size=10,colour=SU_cols("charcoal"),margin=margin(b=4,unit="pt")),
          plot.caption = element_text(hjust = 0,face="italic",size=9,colour=SU_cols("slate"),margin=margin(b=4,unit="pt")),
          legend.text = element_text(size=10,colour=SU_cols("charcoal")),
          legend.title = element_text(face="bold",size=11,colour=SU_cols("charcoal"),margin=margin(b=4,unit="pt")))
}

theme_set(theme_SU())

# Import data/ functions ----

hrg_lookup <-
  tribble(
  ~hrg_3, ~procedure_desc, ~procedure_desc_short, ~procedure_group,
  'BZ3', "Cataract and Lens Procedures",       "Cataract",            "Subchapter BZ – Eyes and Periorbita Procedures and Disorders",
  'BZ4', "Oculoplastics Procedures",           "Oculoplastics",       "Subchapter BZ – Eyes and Periorbita Procedures and Disorders",
  'BZ5', "Orbit and Lacrimal Procedures",      "Orbit and Lacrimal",  "Subchapter BZ – Eyes and Periorbita Procedures and Disorders",
  'BZ6', "Cornea and Sclera Procedures",       "Cornea and Sclera",   "Subchapter BZ – Eyes and Periorbita Procedures and Disorders",
  'BZ7', "Ocular Motility Procedures",         "Ocular Motility",     "Subchapter BZ – Eyes and Periorbita Procedures and Disorders",
  'BZ8', "Vitreous Retinal Procedures",        "Vitreous Retinal",    "Subchapter BZ – Eyes and Periorbita Procedures and Disorders",
  'BZ9', "Glaucoma Procedures",                "Glaucoma",            "Subchapter BZ – Eyes and Periorbita Procedures and Disorders",
  
  'HN1', "Hip Procedures for Non-Trauma",      "Hip",                 "Subchapter HN – Orthopaedic Non-Trauma Procedures",
  'HN2', "Knee Procedures for Non-Trauma",     "Knee",                "Subchapter HN – Orthopaedic Non-Trauma Procedures",
  'HN3', "Foot Procedures for Non-Trauma",     "Foot",                "Subchapter HN – Orthopaedic Non-Trauma Procedures",
  'HN4', "Hand Procedures for Non-Trauma",     "Hand",                "Subchapter HN – Orthopaedic Non-Trauma Procedures",
  'HN5', "Shoulder Procedures for Non-Trauma", "Shoulder",            "Subchapter HN – Orthopaedic Non-Trauma Procedures", 
  'HN6', "Elbow Procedures for Non-Trauma",    "Elbow",               "Subchapter HN – Orthopaedic Non-Trauma Procedures"
  )


```
<hr style="border-top: 2px solid #f9bf07">
# Introduction

The role of the independent sector in the provision of elective care has been brought into the spotlight by rising pressures of the NHS and growing waiting lists. Our analysis aims to contribute to existing work undertaken by the Health Foundation investigating the landscape of [independent sector healthcare services](https://www.health.org.uk/publications/long-reads/waiting-for-nhs-hospital-care-the-role-of-the-independent-sector). 

The Health Foundation found an increasing proportion of care is delivered by the independent sector; this varies by specialitiy and is highest in T&O, GI and ophthalmology. Treatment rates in all specialities are not achieving the 30%-increase target set to address the backlog of care. The Health Foundation suggests uptake in the independent sector is matching the reductions in NHS provided care. It was also found that the independent sector share of provision varies regionally and by deprivation measures. 

We aim to further the above analysis by utilising patient-level inpatient and outpatient data to more closely examine trends in independent sector care. Given the above findings, we will focus on ophthalmology and orthopaedic care, presenting our findings by demographics, geographically and by procedure. 

<hr style="border-top: 2px solid #f9bf07">

# Methods

In an exploratory fashion, we will present the changing market share of independent care in elective ophthalmic and orthopaedic services. We will be guided by our preliminary findings starting by assessing patterns in national data developing on to consider local findings.   

## Data collection 

<div class = "row">
<div class = "col-md-5">
<br>
We gathered acute inpatient and outpatient data via the Secondary Uses Services (SUS) dataset, accessed through the National Commissioning Data Repository (NCDR) platform. To capture the pre-pandemic context, we limit our data to the years 2017 through to 2022. We count completed inpatient admissions or outpatient attendances filtered to only include non-emergency contacts with selected Healthcare Resource Group (HRG) codes. 

The HRG code describes the content of an admission in terms of resource required and is linked to the of commissioning care. We use commissioning codes rather than procedure codes because data completeness is preferable and they offer the suitable level of granularity for data processing and presentation of findings. All HRG codes were updated to their corresponding code in the most recent HRG version. The activity we captured was linked to the following HRG codes (right).

Data was accessed at a national level with local feilds included to indicate the patient's ICS and local authority of residence. Similarly, the low level geography (LSOA) of the patient was included to link to the Index of Multiple Deprivation (IMD) to assign each admission to a deprivation decile (1 = most deprived 10% of the population, 10 = the most affluent 10% of the population).

The SUS data contains only NHS-funded care. We identified the sector in which the care was delivered by assessing the ODS provider type, linked to site (hospital/care centre) in which the care occurred. We compare the rate of NHS-funded care delivered in NHS sites to NHS-funded care delivered in independent sector sites. Limitations of our data dictate that we do not count any privately funded care, regardless of site of delivered.

</div>
  
<div class = "col-md-7">
<br>
***Subchapter BZ – Eyes and Periorbita Procedures and Disorders:***
```{r hrg_codes_eyes_2}
create_dt(
  hrg_lookup %>% 
  filter(str_detect(procedure_group, "Eyes")) %>% 
  select(1,2) %>% 
  rename(HRG = hrg_3, 
         Procedure = procedure_desc)
  )
```

***Subchapter HN – Orthopaedic Non-Trauma Procedures:***
```{r hrg_codes_ortho_2}
create_dt(
  hrg_lookup %>% 
  filter(str_detect(procedure_group, "Orthopaedic")) %>% 
  select(1,2) %>% 
  rename(HRG = hrg_3, 
         Procedure = procedure_desc)
  )
```

</div>
</div>

## Data processing 

Activity data was grouped by demographic, geographic and activity-type variables and counts of episodes, spells (admissions or attendances) and individuals were calculated. Aggregated data was exported from the NCDR SQL server and processed locally in RStudio/RMarkdown. 

For each group of interest we had activity counts for care delivered in both the independent sector (IS) and in NHS clinical settings. We calculated the independent sector market share (%) as follows:

$$
IndependentSectorShare  = ISActivity / (ISActivity + NHSActivity) * 100
$$

Incomplete admissions/attendances, i.e. those without a discharge date, were excluded. Similarly, those without valid identification numbers were excluded.

Upon publication of our findings, the underlying code will be available on [Github](https://github.com/alexanderlawless).

<hr style="border-top: 2px solid #f9bf07">

# Next: Results

National results: <https://alexanderlawless.github.io/Independent_sector_provision/markdown_national_results.html>

Orthopaedic Sub-national results: <https://alexanderlawless.github.io/Independent_sector_provision/markdown_regional_local_results_ophthalmology.html>

Ophthalmic Sub-national results: <https://alexanderlawless.github.io/Independent_sector_provision/markdown_regional_local_results_orthopaedic.html>


<hr style="border-top: 2px solid #f9bf07">


